We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-02-27 paper/paper.tex
agradecimientos

We found that in the best case (Illustris1-Dark) the degree of
asphericity in the observed LG is only expected in $3\pm2$  pairs out
of a sample of $10^4$ isolated pairs. 
This places the LG as a $3\sigma$ outlier.   
The weight to explain this atypical result is not distributed equally
between the MW and M31. 
While M31 presents a fully typical asphericity in the expectations
from LCDM, the MW shows aspherical deviations in plane width and the
major-to-minor axis ratio highly atypical in the  framework of LCDM,
confirming the original hint by \cite{2005A&A...431..517K} and more
recent results by \cite{2015ApJ...815...19P}. 
We estimated that with the M31 between $37\%$ and $58\%$ of the pairs
pairs show aspherical characteristics larger than M31, while
this fraction drops to less than $1\%$ for the MW. 
These fractions are robust to changes in the numerical simulations
and, the criteria used to define the pairs and the methods used to
estimate the parameters of the multivariate normal distribution.

@@

The LG atypicality should be seen as an opportunity to constrain in
high detail the environment that allowed such a pattern to emerge. 
Although local correlations between LG assembly, pair kinematics, halo shapes and
satellite distributions are expected in LCDM
\citep{2011MNRAS.417.1434F,2014MNRAS.443.1090F,2015ApJ...799...45F,2015MNRAS.452.1052L},
we claim that detailed studies of the satellite asphericity as a function
of halo mass and cosmic web environment are still missing to
understand what fearures in the initial conditions of our Local Group
are responsible for the assembly extreme features observed in its satellite
distribution.

\section*{Acknowledgements} 
We acknowledge Universidad de los Andes and COLCIENCIAS (Project -
XXX) for the computing resources used in this project.

>>>>

We found that in the best case (Illustris1-Dark) the degree of
asphericity in the observed LG is only expected in $3\pm2$  pairs out
of a sample of $10^4$ isolated pairs. 
This places the LG as a $3\sigma$ outlier.   
The weight to explain this atypical result is not distributed equally
between the MW and M31. 
While M31 presents a fully typical asphericity in the expectations
from LCDM, the MW shows aspherical deviations in plane width and the
major-to-minor axis ratio highly atypical in the  framework of LCDM,
confirming the original hint by \cite{2005A&A...431..517K} and more
recent results by \cite{2015ApJ...815...19P}. 
We estimated that with the M31 between $37\%$ and $58\%$ of the pairs
pairs show aspherical characteristics larger than M31, while
this fraction drops to less than $1\%$ for the MW. 
These fractions are robust to changes in the numerical simulations, 
the criteria used to define the pairs and the methods used to
estimate the parameters of the multivariate normal distribution.

@@

The LG atypicality should be seen as an opportunity to constrain in
high detail the environment that allowed such a pattern to emerge. 
Although broad correlations between LG assembly, pair kinematics, halo shapes and
satellite distributions are expected in LCDM
\citep{2011MNRAS.417.1434F,2014MNRAS.443.1090F,2015ApJ...799...45F,2015MNRAS.452.1052L},
we claim that detailed studies of the satellite asphericity as a function
of halo mass and cosmic web environment are still missing to
understand what fearures in the initial conditions of our LG
are responsible for the extreme features observed in its satellite
distribution.

\section*{Acknowledgements} 
We acknowledge financial support from: Universidad de Los Andes
(Colombia); COLCIENCIAS (c\'odigo 120471250459, contrato FP44842-287-
2016); the European Union’s Horizon 2020 Research and Innovation
Programme under the Marie Sk\l{}odowska-Curie grant agreement No 73437
(LACEGAL).
